---
title: "pre class 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Background

The exponential distribution is defined by its cumulative distribution function
\(F(x) = 1-e^{-\lambda x}\)

The R function ***rexp()*** generates random variables with an exponential distribution. For example 
<center><strong>rexp(n=10, rate=5)</strong> </center>

results in 10 exponentially distributed numbers with a rate \(\lambda=5\). If you leave out the 5 and just have
<center><strong>rexp(n=10) </strong></center>
then this results in 10 exponentially distributed numbers with a rate \(\lambda=1\), this is also referred to as the "standard exponential distribution". 

### Part 1

1. Generate 200 random values from the standard exponential distribution and store them in a vector `exp.draws.1`.  Find the mean and standard deviation of `exp.draws.1`.

```{r q1}
exp.draws.1 = rexp(n = 200) 
mean.draws.1 = mean(exp.draws.1)
sd.draws.1 = sd(exp.draws.1)
print(paste("The mean of exp.draws.1 is", mean.draws.1))
print(paste("The standard deviation of exp.draws.1 is", sd.draws.1))
```

2. Repeat, but change the rate to 0.2, 5, 7.3 and 10, storing the results in vectors called  `exp.draws.0.2`,  `exp.draws.5`,  `exp.draws.7.3` and  `exp.draws.10`. 

```{r q2}
exp.draws.0.2 = rexp(n = 200, rate = 0.2)
exp.draws.5 = rexp(n = 200, rate = 5)
exp.draws.7.3= rexp(n = 200, rate = 7.3)
exp.draws.10 = rexp(n = 200, rate = 10)

mean.draws.0.2 = mean(exp.draws.0.2)
sd.draws.0.2 = sd(exp.draws.0.2)
print(paste("The mean of exp.draws.0.2 is", mean.draws.0.2, "The standard deviation of exp.draws.0.2 is", sd.draws.0.2))

mean.draws.5 = mean(exp.draws.5)
sd.draws.5 = sd(exp.draws.5)
print(paste("The mean of exp.draws.5 is", mean.draws.5, "The standard deviation of exp.draws.5 is", sd.draws.5))

mean.draws.7.3 = mean(exp.draws.7.3)
sd.draws.7.3 = sd(exp.draws.7.3)
print(paste("The mean of exp.draws.7.3 is", mean.draws.7.3, "The standard deviation of exp.draws.7.3 is", sd.draws.7.3))

mean.draws.10 = mean(exp.draws.10)
sd.draws.10 = sd(exp.draws.10)
print(paste("The mean of exp.draws.10 is", mean.draws.10, "The standard deviation of exp.draws.10 is", sd.draws.10))
```

3. The function `plot()` is the generic function in R for the visual display of data. `hist()` is a function that takes in and bins data as a side effect. To use this function, we must first specify what we'd like to plot.
    a. Use the `hist()` function to produce a histogram of your standard exponential distribution. 
    b. Use `plot()` with this vector to display the random values from your standard distribution in order.
    c. Now, use `plot()` with two arguments -- any two of your other stored random value vectors -- to create a scatterplot of the two vectors against each other.

```{r q3}
hist(exp.draws.1)
plot(exp.draws.1, main = "Plots for exp.draw.1")
plot(exp.draws.0.2, exp.draws.5, main = "Plots for exp.draws.0.2 and exp.draws.5")
```
```{Explanations of the plots in question 3}
a. The histogram of the exp.draws.1 has the similar trend with the density function curve of standard exponential distribution. Large frequency is located near zero and decreases exponentially with the increase of the exp.draws.1.
b. From the plot of exp.draws.1, we can see that the dots are densely distributed near the value of zero, while loosely distributed far way from the value of zero, which is coincide with the frequency from the histogram.
c. In this plot, I compared draws.exp.0.2 and draws.xp.5. The dots are densely located near origin and loosely located far away from the origin. This means it is more likely happen that both exp.draws.0.2 and exp.draws.5 take small values around zero. Besides, we can also see some dots near x axis and y axis, which means it is likely that at least one of the two random numbers is small.
``` 


4. We'd now like to compare the properties of each of our vectors. Begin by creating a vector of the means of each of our five distributions in the order we created them and saving this to a variable name of your choice. Using this and other similar vectors, create the following scatterplots and explain in words what is going on:
    a. The five means versus the five rates used to generate the distribution.
    b. The standard deviations versus the rates.
    c. The means versus the standard deviations.

For each plot, explain in words what's going on.

```{r q4}
mean.exp.all = c(mean.draws.1, mean.draws.0.2, mean.draws.5, mean.draws.7.3, mean.draws.10)
rates.exp.all = c(1, 0.2, 5, 7.3, 10)
sd.exp.all = c(sd.draws.1, sd.draws.0.2, sd.draws.5, sd.draws.7.3, sd.draws.10)

plot(mean.exp.all, rates.exp.all, main = "Five means vs five rates")
plot(sd.exp.all, rates.exp.all, main = "Five sd vs five rates")
plot(mean.exp.all, sd.exp.all, main = "Five means vs five sd")
```
```{Explanations of Plots in question 4}
a. From the plots of mean vs. rates, we can see that large rates are corresponding with small means and vice versa. The rates and the means have negative realationship.
b. The plot of standard deviations and the plot of mean vs.rates are similar, which means the standard deviation and the mean of each random exponential distributions are numerical similar. And the standard deviation also has the negative relationship with the rates.
c. In this plot, we compare means and the standard deviations of the random exponential numbers. All of the dots located near the line of y=x, which means the means and the standard deviations are similar for each rates. The larger the rates, the smaller the means and standard deviations.

```

### Part II (PHP 2560 Only)


5. R's capacity for data and computation is large to what was available 10 years ago. 
    a. To show this, generate 1.1 million numbers from the standard exponential distribution and store them in a vector called `big.exp.draws.1`. Calculate the mean and standard deviation.
```{r q5 part_a}
big.exp.draws.1 = rexp(n = 1100000)
mean.big.draws.1 = mean(big.exp.draws.1)
sd.big.draws.1 = sd(big.exp.draws.1)
print(paste("The mean of the big.exp.draws.1 is ", mean.big.draws.1))
print(paste("The standard deviations of the big.exp.draws.1 is ", sd.big.draws.1))
```
    
    b. Plot a histogram of `big.exp.draws.1`.  Does it match the function \(1-e^{-x}\)?  Should it? 
```{r q5 part_b}
hist(big.exp.draws.1)
```
```{q5 part_b}
This matches the exponential distribution because the number of random numbers is pretty large.
```

    c. Find the mean of all of the entries in `big.exp.draws.1` which are strictly greater than 1. You may need to first create a new vector to identify which elements satisfy this.
```{r q5 part_c}
larger.than.1 = big.exp.draws.1[big.exp.draws.1>1]
mean.larger.than.1 = mean(larger.than.1)
print(paste("The mean of all of the entries which are strictly larger than 1 is ", mean.larger.than.1))
```
    
    d. Create a matrix, `big.exp.draws.1.mat`, containing the the values in 
`big.exp.draws.1`, with 1100 rows and 1000 columns. Use this matrix as the input to the `hist()` function and save the result to a variable of your choice. What happens to your data?
```{r q5 part_d}
big.exp.draws.1.mat = matrix(big.exp.draws.1, nrow = 1100, ncol = 1000)
hist(big.exp.draws.1.mat)
```
```{}
The plot of 5(d) is the same as 5(a).
```

    e. Calculate the mean of the 371st column of `big.exp.draws.1.mat`.
```{r q5 part_e}
mean.371 = mean(big.exp.draws.1.mat[,371])
print(paste("The mean of the 371st column of big.exp.draws.1.mat is ", mean.371))
```

    f. Now, find the means of all 1000 columns of `big.exp.draws.1.mat` simultaneously. Plot the histogram of column means.  Explain why its shape does not match the histogram in problem 5b).
```{r q5 part_f}
mean.1000 = apply(big.exp.draws.1.mat, 2, mean)
hist(mean.1000)
```
```{q5 part_f}
The curve of mean of all 1000 columns looks like a normal distribution with the center of 1. This looks different with 5(a) because it dipicts the frequency distribution of the mean of the random numbers, not the random numbers themselves. The means are near to 1/rate, which is 1 in this case. 
```

    
    