# pre-class


Make sure you commit this often with meaningful messages. 


### Getting Started


We will work with the dataset called [gapminder](https://github.com/jennybc/gapminder), this is a cleaned up version from [Gapminder Data](http://www.gapminder.org/data/). Gapminder contains a lot of great data on all of the nations of the world. We first need to install the gapminder package in R. 

```
install.packages("gapminder")
```


```{r}
library(dplyr)
library(gapminder)
gapminder
class(gapminder)
```



### Pre-Class Problems

Use **dplyr** functions to address the following questions:

1. How many unique countries are represented per continent?
```{r}
gapminder %>%
  group_by(continent) %>%
  summarise(n_distinct(country))
```

2. Which European nation had the lowest GDP per capita in 1997? 
```{r}
gapminder %>%
  filter(year == 1997 & continent == 'Europe') %>%
  arrange(gdpPercap) %>%
  head(1)
```

3. According to the data available, what was the average life expectancy across each continent in the 1980s?
```{r}
gapminder %>%
  filter(year >=1980 & year <= 1989) %>% 
  group_by(continent) %>%
  summarise(meanExp = mean(lifeExp))
```

4. What 5 countries have the highest total GDP over all years combined?
```{r}
gapminder %>%
  group_by(country) %>%
  summarise(total_gdp = sum(gdpPercap)) %>%
  arrange(desc(total_gdp)) %>%
  head(5)
```

5. What countries and years had life expectancies of _at least_ 80 years? _N.b. only output the columns of interest: country, life expectancy and year (in that order)._
```{r}
gapminder %>%
  filter(lifeExp >= 80) %>%
  select(country, year, lifeExp)
```

6. What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?
```{r}
gapminder %>%
  group_by(country) %>%
  summarise(corr = cor(lifeExp, gdpPercap)) %>%
  arrange(desc(corr)) %>%
  head(10)
```

7. Which combinations of continent (besides Asia) and year have the highest average population across all countries? _N.b. your output should include all results sorted by highest average population_. With what you already know, this one may stump you. See [this Q&A](http://stackoverflow.com/q/27207963/654296) for how to `ungroup` before `arrange`ing. This also [behaves differently in more recent versions of dplyr](https://github.com/hadley/dplyr/releases/tag/v0.5.0).
```{r}
unique(gapminder$continent)
gapminder %>%
  filter(continent != "Asia") %>%
  group_by(continent, year) %>%
  summarise(ave_pop = sum(pop)/n_distinct(country)) %>%
  ungroup() %>%
  arrange(desc(ave_pop))
```

8. Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data? 
```{r}
gapminder %>%
  group_by(country) %>%
  summarise(sd_pop = sd(pop)) %>%
  arrange(sd_pop) %>%
  head(3)
```

9. Subset **gm** to only include observations from 1992 and store the results as **gm1992**. What kind of object is this?
```{r}
gm1992 = gapminder[gapminder$year >= 1992, ]
gm1992
class(gm1992)
```

10. Which observations indicate that the population of a country has *decreased* from the previous year **and** the life expectancy has *increased* from the previous year? See [the vignette on window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html).
```{r}
gapminder %>%
  arrange(country, year) %>%
  filter(pop < lag(pop) & lifeExp > lag(lifeExp) & year > lag(year)) 

```
