# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}
check__string_identical <- function(string, prefix) { ## The function is to check whether prefix is equals to the first part of string
  substr(string, 1, nchar(prefix)) == prefix
  print(prefix)
}
delete_end <- function(x) { ## Remove the last one item of a vector if the vector contains more than one elements
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
repeat_exact_length <- function(x, y) { ## Repeat y the length of x times
  rep(y, length.out = length(x))
}
```


2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?

*rnorm() is the function to generate random numbers from one-dimension normal distribution, with default value mu = 0 and sigma = 1. We need to set how many numbers we want to generate. We can also change mu and sigma as well
mvrnorm() is the function to generate one or more samples from the specific multivariate distribution, with default value n = 1, tol = 1e-6.We need to specify the mu and sigma, we can change number of random vectors as well. *

```{r}
## Use mvrnorm to generate rnorm

library(MASS)
generate_rnorm = function(n, mu = 0, Sigma = 1) {
  stopifnot(length(mu) == 1 & ncol(sigma) == 1 & nrow(Sigma) == 1)
  as.vector(mvrnorm(n, mu, Sigma))
}

generate_rnorm(3)

```


3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 

```{r}
data(mtcars)
lapply(mtcars, function(x) sd(x,na.rm = TRUE)/mean(x, na.rm = TRUE))

```

4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.
    b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)
    
```{r}
sd_vapply = function(x) {
  stopifnot(is.data.frame(x))
  vapply(x, function(x,na.rm = T) sd(x), numeric(1))
}
#sd_vapply(mtcars)

sd_vapply = function(x) {
  stopifnot(is.data.frame(x))
  judge = vapply(x, is.numeric, logical(1))
  vapply(x[judge], function(x,na.rm = T) sd(x), numeric(1))
  
}
#new_data = data.frame("SN" = 1:2, "Age" = c(21,15), "Name" = c("John","Dora"))
#sd_vapply(new_data)
```
